### 2018-04-20 resistingrhythm/analysis/ca_analysis_v2.Rmd
Added new Ca analysis Rmd, adj col names to match the new format

---
title: "Ca exp analysis -- 4/18/2018"
output: html_notebook
---

```{r}
library(tidyverse)
library(gridExtra)

data_path <- "~/Code/resistingrhythm/data/"
```

```{r, message=FALSE}
exp_code <- 1
osc_codes <- c(1, 3, 5)
osc_rates <- c(1, 3, 5)
stim_codes <- c(0)

# Load exp 1 results for comparison

results <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$Ca_target <- rep("0.030", nrow(df))

    results <- rbind(results, df)
  }
}

# Rename Ca to match new format
results %>% 
  rename(Ca_obs_target = Ca_k) %>% 
  mutate(Ca_obs_ref = Ca_obs_target) -> results

# ------------------------------------------------------------------
# exp_code <- 2
# ca_values <- c("0.032", "0.034", "0.036", "0.038")

exp_code <- 7
ca_values <- c("0.032", "0.034", "0.036")
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    for(ca in ca_values){
      o_c <- osc_codes[i]
      o_r <- osc_rates[i]

      # Generate name
      file_name <- paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         "_Ca", ca,
                         ".csv", sep="")

      if (file.exists(file_name)){
        # Load
        df <- read_csv(file_name)

        # Add metadata
        df$osc_code <- rep(o_c, nrow(df))
        df$osc_rate <- rep(o_r, nrow(df))
        df$stim_code <- rep(s_c, nrow(df))

        results <- rbind(results, df)  
      }
    }
  }
}

rm(df, exp_code, ca_values, osc_codes, osc_rates, file_name, i, o_c, o_r, s_c, stim_codes, ca)
```

# Osc rate 
## By variance

```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=kappa_coord, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Kappa", x="Osc. rate (Hz)") +
  theme_classic() 
```
```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_var, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Variance", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By error

```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_error, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Error", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By rate

```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Firing rate (Hz)", x="Rate (Hz") +
  theme_classic() 
```

```{r, fig.width=2.0, fig.height=1.0}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_ref - rate_k, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_hline(yintercept=0, size=0.5) +
  labs(y="Rate diff. (Hz)", x="Osc. rate (Hz)") +
  theme_classic() 
```

# Error versus var

```{r, fig.width=2.2, fig.height=0.8}
results %>% 
  ggplot(aes(x=mse_error, y=kappa_coord, color=Ca_target)) +
  geom_point(alpha=0.5, size=0.2) + 
  theme_classic() + 
  labs(y="Kappa", x="Error") +
  facet_grid(.~osc_rate)
```

```{r, fig.width=1.8, fig.height=0.8}
results %>% 
  ggplot(aes(x=mse_error, y=mse_var, color=Ca_target)) +
  geom_point(alpha=0.5, size=0.2) + 
  theme_classic() + 
  labs(y="Variance", x="Error") +
  facet_grid(.~osc_rate)
```