### 2018-04-20 resistingrhythm/analysis/burst_analysis.Rmd
Adding control_analysis and burst_analysis

---
title: "General burst analysis -- 4/18/2018"
output: html_notebook
---

```{r}
library(tidyverse)
library(gridExtra)

data_path <- "~/Code/resistingrhythm/data/"
```

```{r, message=FALSE}
exp_code <- 3
osc_codes <- c(0, 1, 3, 5)
osc_rates <- c(0.5, 1, 3, 5)
stim_codes <- c(0, 1, 2)

results <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))

    results <- rbind(results, df)
  }
}
```

# Osc rate 
## By variance

```{r, fig.width=1.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=kappa_coord, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Kappa", x="Osc. rate (Hz)") +
  theme_classic() 
```

```{r, fig.width=1.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_var, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Variance", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By error

```{r, fig.width=1.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_error, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Error", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By rate

```{r, fig.width=1.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Mod. rate (Hz)", x="Osc. rate (Hz)") +
  theme_classic() 
```

```{r, fig.width=1.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k - rate_ref, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_hline(yintercept=0, size=0.5) +
  labs(y="Rate diff. (Hz)", x="Osc. rate (Hz)") +
  theme_classic() 
```

# Error versus var

```{r, fig.width=1.8, fig.height=0.8}
results %>% 
  ggplot(aes(x=mse_error, y=kappa_coord)) +
  geom_point(size=0.4, alpha=0.2) + 
  theme_classic() + 
  labs(y="Kappa", x="Error") +
  facet_grid(.~osc_rate)
```

```{r, fig.width=1.8, fig.height=0.8}
results %>% 
  ggplot(aes(x=mse_error, y=mse_var)) +
  geom_point(size=0.4, alpha=0.2) + 
  theme_classic() + 
  labs(y="Variance", x="Error") +
  facet_grid(.~osc_rate)
```