### 2018-04-28 resistingrhythm/analysis/burst_analysis.Rmd
Rmd vis - tweaks to nearly everything

---
title: "General burst analysis -- 4/18/2018"
output: html_notebook
---

```{r}
library(tidyverse)
library(gridExtra)

data_path <- "~/Code/resistingrhythm/data/"
```

```{r, message=FALSE}

exp_code <- 16
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130, 135, 140, 145, 150, 155, 160)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0)
stim_codes <- c(0, 1)

results <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))

    results <- rbind(results, df)
  }
}

# exp_code <- 3
# osc_codes <- c(0, 1, 3, 5)
# osc_rates <- c(0.5, 1, 3, 5)
# stim_codes <- c(0, 1, 2)
# 
# results0 <- NULL
# for(i in 1:length(osc_codes)){
#   for(s_c in stim_codes){
#     o_c <- osc_codes[i]
#     o_r <- osc_rates[i]
#     
#     # Load the csv
#     df <- read_csv(paste(data_path, 
#                          "exp", as.character(exp_code), 
#                          "_stim", as.character(s_c), 
#                          "_burst", as.character(o_c),
#                          ".csv", sep=""))
#     
#     # Add metadata
#     df$osc_code <- rep(o_c, nrow(df))
#     df$osc_rate <- rep(o_r, nrow(df))
#     df$stim_code <- rep(s_c, nrow(df))
#     
#     results0 <- rbind(results0, df)
#   }
# }
# 
# 
# exp_code <- 14
# osc_codes <- c(7, 8, 9, 10, 11, 12, 13, 14, 15)
# osc_rates <- c(0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45)
# stim_codes <- c(0, 1, 2)
# 
# results1 <- NULL
# for(i in 1:length(osc_codes)){
#   for(s_c in stim_codes){
#     o_c <- osc_codes[i]
#     o_r <- osc_rates[i]
#     
#     # Load the csv
#     df <- read_csv(paste(data_path, 
#                          "exp", as.character(exp_code), 
#                          "_stim", as.character(s_c), 
#                          "_burst", as.character(o_c),
#                          ".csv", sep=""))
#     
#     # Add metadata
#     df$osc_code <- rep(o_c, nrow(df))
#     df$osc_rate <- rep(o_r, nrow(df))
#     df$stim_code <- rep(s_c, nrow(df))
#     
#     results1 <- rbind(results1, df)
#   }
# }
# 
# results0 %>% 
#   select(k, osc_rate, osc_code, kappa_error, kappa_coord, mse_error, 
#          mse_var, abs_error, abs_var, rate_k, rate_ref, stim_code) -> results0
# results1 %>% 
#   select(k, osc_rate, osc_code, kappa_error, kappa_coord, mse_error, mse_var, abs_error, abs_var, rate_k, rate_ref, stim_code) -> results1
# results <- rbind(results0, results1)

results %>% 
  filter(kappa_coord > 0) -> results
```

# Osc rate 
## By variance

```{r, fig.width=2.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=kappa_coord, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  # geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Kappa", x="Osc. rate (Hz)") +
  theme_classic() 
```

```{r, fig.width=2.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_var, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Variance", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By error

```{r, fig.width=3.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_error, group=osc_rate)) +
  # geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Error", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By rate

```{r, fig.width=2.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Mod. rate (Hz)", x="Osc. rate (Hz)") +
  theme_classic() 
```

```{r, fig.width=2.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k - rate_ref, group=osc_rate)) +
  geom_jitter(alpha=0.1, size=0.1) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_hline(yintercept=0, size=0.5) +
  labs(y="Rate diff. (Hz)", x="Osc. rate (Hz)") +
  theme_classic() 
```

# Error versus var

```{r, fig.width=2.8, fig.height=0.8}
results %>% 
  ggplot(aes(x=mse_error, y=kappa_coord)) +
  geom_point(size=0.4, alpha=0.2) + 
  geom_smooth(method='lm', color="red", alpha=0.4, se=FALSE, size=0.1) +
  theme_classic() + 
  labs(y="Kappa", x="Error") +
  facet_grid(.~osc_rate)
```

```{r, fig.width=2.8, fig.height=0.8}
results %>% 
  ggplot(aes(x=mse_error, y=mse_var)) +
  geom_point(size=0.4, alpha=0.2) + 
  geom_smooth(method='lm', color="red", alpha=0.4, se=FALSE, size=0.1) +
  theme_classic() + 
  labs(y="Variance", x="Error") +
  facet_grid(.~osc_rate)
```