### 2018-04-28 resistingrhythm/analysis/ca_analysis_v2.Rmd

---
title: "Ca exp analysis -- 4/18/2018"
output: html_notebook
---

```{r}
library(tidyverse)
library(gridExtra)

data_path <- "~/Code/resistingrhythm/data/"
```

```{r, message=FALSE}
exp_code <- 19
osc_codes <- c(101, 103, 105, 115, 125, 135, 145, 155)
osc_rates <- c(0.1, 0.3, 0.5, 1.5, 2.5, 3.5, 4.5, 5.5)
stim_codes <- c(0, 1)

ca_targets <- c(0.026, 0.028, 0.032, 0.034)

results <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    for(ca in ca_targets){
      o_c <- osc_codes[i]
      o_r <- osc_rates[i]

      # Load the csv
      df <- read_csv(paste(data_path,
                           "exp", as.character(exp_code),
                           "_stim", as.character(s_c),
                           "_osc", as.character(o_c),
                           "_ca", as.character(ca),
                           ".csv", sep=""))

      # Add metadata
      df$osc_code <- rep(o_c, nrow(df))
      df$osc_rate <- rep(o_r, nrow(df))
      df$stim_code <- rep(s_c, nrow(df))
      df$Ca_target <- rep(ca, nrow(df))

      results <- rbind(results, df)
    }
  }
}

# exp_code <- 1
# osc_codes <- c(1, 3, 5)
# osc_rates <- c(1, 3, 5)
# stim_codes <- c(0)
# 
# # Load exp 1 results for comparison
# 
# results <- NULL
# for(i in 1:length(osc_codes)){
#   for(s_c in stim_codes){
#     o_c <- osc_codes[i]
#     o_r <- osc_rates[i]
#     
#     # Load the csv
#     df <- read_csv(paste(data_path, 
#                          "exp", as.character(exp_code), 
#                          "_stim", as.character(s_c), 
#                          "_osc", as.character(o_c),
#                          ".csv", sep=""))
#     
#     # Add metadata
#     df$osc_code <- rep(o_c, nrow(df))
#     df$osc_rate <- rep(o_r, nrow(df))
#     df$stim_code <- rep(s_c, nrow(df))
#     df$Ca_target <- rep("0.030", nrow(df))
#     
#     results <- rbind(results, df)
#   }
# }
# 
# # Rename Ca to match new format
# results %>% 
#   rename(Ca_obs_target = Ca_k) %>% 
#   mutate(Ca_obs_ref = Ca_obs_target) -> results
# 
# # ------------------------------------------------------------------
# # exp_code <- 2
# # ca_values <- c("0.032", "0.034", "0.036", "0.038")
# 
# exp_code <- 7
# ca_values <- c("0.032", "0.034", "0.036")
# for(i in 1:length(osc_codes)){
#   for(s_c in stim_codes){
#     for(ca in ca_values){
#       o_c <- osc_codes[i]
#       o_r <- osc_rates[i]
#       
#       # Generate name
#       file_name <- paste(data_path, 
#                          "exp", as.character(exp_code), 
#                          "_stim", as.character(s_c), 
#                          "_osc", as.character(o_c),
#                          "_Ca", ca,
#                          ".csv", sep="")
#     
#       if (file.exists(file_name)){
#         # Load
#         df <- read_csv(file_name)
# 
#         # Add metadata
#         df$osc_code <- rep(o_c, nrow(df))
#         df$osc_rate <- rep(o_r, nrow(df))
#         df$stim_code <- rep(s_c, nrow(df))
#         
#         results <- rbind(results, df)  
#       }
#     }
#   }
# }

rm(df, exp_code, ca_values, osc_codes, osc_rates, file_name, i, o_c, o_r, s_c, stim_codes, ca)
```

# Osc rate 
## By variance

```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=kappa_coord, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  # geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_jitter(size=.1) +
  facet_grid(.~Ca_target) +
  labs(y="Kappa", x="Osc. rate (Hz)") +
  theme_classic() 
```
```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_var, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  # geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_jitter(size=.1) +
  facet_grid(.~Ca_target) +
  labs(y="Variance", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By error

```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_error/1e3, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  # geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_jitter(size=.1) +
  facet_grid(.~Ca_target) +
  labs(y="Error", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By rate

```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  # geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_jitter(size=.1) +
  facet_grid(.~Ca_target) +
  labs(y="Mod. rate (Hz)", x="Rate (Hz") +
  theme_classic() 
```

```{r, fig.width=2.0, fig.height=1.0}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k - rate_ref, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  # geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_jitter(size=.1) +
  facet_grid(.~Ca_target) +
  labs(y="Delta rate (Hz)", x="Rate (Hz)") +
  theme_classic() 
```

# Error versus var

```{r, fig.width=2, fig.height=2}
results %>% 
  ggplot(aes(x=mse_error/1e3, y=mse_var, 
             group=interaction(osc_rate, Ca_target),
             color=factor(Ca_target))) +
  # geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  geom_jitter(size=.1) +
  facet_grid(Ca_target~.) +
  labs(y="Variance", x="Error") +
  theme_classic() 
```