### 2018-04-28 resistingrhythm/analysis/step_analysis.Rmd

---
title: "Ca exp analysis -- 4/18/2018"
output: html_notebook
---

```{r}
library(tidyverse)
library(gridExtra)

data_path <- "~/Code/resistingrhythm/data/"
```

```{r, message=FALSE}
exp_code <- 17
osc_codes <- c(200, 201, 202, 203, 204, 205)
osc_rates <- c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5)

step_codes <- c(0, 1, 2, 3)
step_currents <- c(0.0, -0.5e-6, -0.75e-6, -1.0e-6)

stim_codes <- c(0, 1)

results <- NULL
for(i in 1:length(osc_codes)){
  for(j in 1:length(step_codes)){

    for(s_c in stim_codes){
      o_c <- osc_codes[i]
      o_r <- osc_rates[i]

      stp <- step_codes[j]
      I <- step_currents[j]

      # Load the csv
      df <- read_csv(paste(data_path, 
                           "exp", as.character(exp_code), 
                           "_stim", as.character(s_c), 
                           "_burst", as.character(o_c),
                           "_step", as.character(stp),
                           ".csv", sep=""))

      # Add metadata
      df$osc_code <- rep(o_c, nrow(df))
      df$osc_rate <- rep(o_r, nrow(df))
      df$stim_code <- rep(s_c, nrow(df))

      df$step_code <- rep(stp, nrow(df))
      df$step_current <- rep(I, nrow(df))

      results <- rbind(results, df)
    }
  }
}

results %>% 
  filter(kappa_coord > 0) -> results

rm(df, exp_code, step_codes, step_currents, osc_codes, osc_rates, stp, I, j, i, o_c, o_r, s_c, stim_codes)
```

# Osc rate 
## By variance

```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=kappa_coord, 
             group=interaction(osc_rate, step_current),
             color=factor(step_current))) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Kappa", x="Osc. rate (Hz)") +
  theme_classic() 
```
```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_var, 
             group=interaction(osc_rate, step_current),
             color=factor(step_current))) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Variance", x="Osc. rate (Hz)") +
  theme_classic() 
```

## By error

```{r, fig.width=2.0, fig.height=1}
results %>% 
  ggplot(aes(x=osc_rate, y=mse_error, 
             group=interaction(osc_rate, step_current),
             color=factor(step_current))) +
  geom_boxplot(size=0.3, outlier.colour = NA, alpha=0.5) +
  labs(y="Error", x="Osc. rate (Hz)") +
  theme_classic() 
```

# Var vesus error

```{r, fig.width=3.0, fig.height=0.8}
results %>% 
  ggplot(aes(x=mse_error, y=mse_var, 
             color=factor(step_current))) +
  geom_point(size=0.1) +
  labs(y="Variance", x="Error") +
  facet_grid(.~step_current) +
  theme_classic() 
```