### 2018-11-17 resistingrhythm/analysis/exp229.Rmd
exp229 analysis intial commit

---
title: "Analysis : exp229 analysis"
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(grid)
library(ggpubr)
library(cowplot)

data_path <- "~/Code/resistingrhythm/data/"
```

# Load all data
```{r, message=FALSE}
# -----------------------------------------------------------------------
# 16: Burst E, with H
tau_h <- 4.0
exp_code <- 229
run_time_codes <- c('1.0', '2.0', '3.0', '4.0', '5.0', '6.0', '7.0', '8.0')
onset_time_codes <- c('0.5', '1.5', '2.5', '3.5', '4.5', '5.5', '6.5', '7.5')
freq_codes <- c('8', '12', '16', '20', '24', '28')
osc_rate_codes <- c('0', '2', '4', '6')

results229 <- NULL
for(i in 1:length(run_time_codes)){
  for(f_c in freq_codes){
    for(o_r in osc_rate_codes){
      r_t <- run_time_codes[i]
      o_t <- onset_time_codes[i]

      # Load the csv
      df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_T", r_t, 
                         "_t", o_t,
                         "_r", o_r,
                         "_f", f_c,
                         ".csv", sep=""))  
      # Add metadata
      df$run_time <- rep(r_t, nrow(df))
      df$frac_tau_h <- rep(as.character(as.numeric(r_t)/tau_h, nrow(df)))
      df$onset_time <- rep(o_t, nrow(df))
      df$osc_rate <- rep(o_r, nrow(df))
      df$freq <- rep(f_c, nrow(df))
      df$synapse <- rep("Excitatory", nrow(df))
      results229 <- rbind(results229, df)
    }
  }
}

results229$freq <- factor(results229$freq, levels = freq_codes)
results229$run_time <- factor(results229$run_time, levels = run_time_codes)
results229$osc_rate <- factor(results229$osc_rate, levels = osc_rate_codes)
```

```{r, fig.width=1.5, fig.height=1.2}
null_rate <- mean(filter(results229, osc_rate == "0")$rate_k)
results229 %>% 
  filter(osc_rate != "0") %>%
  ggplot(aes(x=frac_tau_h, y=rate_k)) +
    geom_boxplot() +
    theme_classic2() +
    scale_color_brewer(palette="PuOr", direction=-1) +
    labs(x="Frac. tau_h", y="Pop. rate") +
    theme(legend.position = "none")
```

```{r}
results229 %>% 
  filter(osc_rate == "2") %>% 
  ggplot(aes(x=freq, y=rate_k, color=frac_tau_h)) +
  geom_boxplot() +
  theme_classic2() +
  scale_color_brewer(palette="BrBG", direction=-1) +
  labs(x="Frequency (Hz)", y="Pop. rate") 
```

```{r}
results229 %>% 
  filter(osc_rate != "0") %>% 
  ggplot(aes(x=osc_rate, y=rate_k, color=frac_tau_h)) +
  geom_boxplot() +
  theme_classic2() +
  scale_color_brewer(palette="BrBG", direction=-1) +
  labs(x="Osc. rate", y="Pop. rate") 
```