### 2018-12-08 resistingrhythm/analysis/exp231.Rmd

---
title: "Analysis : exp231 analysis"
output: html_notebook
---

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)
library(readr)
library(gridExtra)
library(grid)
library(ggpubr)
library(cowplot)

data_path <- "~/Code/resistingrhythm/data/"
```

# Load all data
```{r, message=FALSE}
# 231
tau_h <- 600
exp_code <- 231
n_cycles <- c("4", "8", "480")

results231 <- NULL
for(i in 1:length(n_cycles)){
  n_c <- n_cycles[i]

  # Load the csv
  df <- read_csv(paste(data_path, 
                       "exp", as.character(exp_code), 
                       "_c", n_c, 
                       ".csv", sep=""))  
  # Add metadata
  df$n_cycles <- rep(n_c, nrow(df))
  df$frac_tau_h <- rep(as.character(
    (
      round(as.numeric(n_c)*(1/8)/tau_h, digits=3)
    ), nrow(df)
    )
  )
  df$osc_rate <- rep("6", nrow(df))
  df$freq <- rep("b", nrow(df))
  df$synapse <- rep("Excitatory", nrow(df))
  df$homeo <- rep("Homeostasis", nrow(df))
  if(n_c <= 8){
    df$mode <- rep("Burst", nrow(df))
  } else {
    df$mode <- rep("Tonic", nrow(df))
  }

  results231 <- rbind(results231, df)
}

exp_code <- 232
results232 <- NULL
for(i in 1:length(n_cycles)){
  n_c <- n_cycles[i]

  # Load the csv
  df <- read_csv(paste(data_path, 
                       "exp", as.character(exp_code), 
                       "_c", n_c, 
                       ".csv", sep=""))  
  # Add metadata
  df$n_cycles <- rep(n_c, nrow(df))
  df$frac_tau_h <- rep(as.character(
    (
      round(as.numeric(n_c)*(1/8)/tau_h, digits=3)
    ), nrow(df)
    )
  )
  df$osc_rate <- rep("6", nrow(df))
  df$freq <- rep("b", nrow(df))
  df$synapse <- rep("Excitatory", nrow(df))
  df$homeo <- rep("No homeostasis", nrow(df))
  if(n_c <= 8){
    df$mode <- rep("Burst", nrow(df))
  } else {
    df$mode <- rep("Tonic", nrow(df))
  }

  results232 <- rbind(results232, df)
}

rm(df, n_c, i, exp_code)
```

```{r, fig.width=2.2, fig.height=1.1}
rbind(results231, results232) %>% 
  ggplot(aes(x=frac_tau_h, y=rate_k, color=homeo, group=interaction(homeo, frac_tau_h))) +
  geom_boxplot(size=0.6) +
  scale_color_manual(values=c("black", "grey"), name="") 
```

```{r, fig.width=2.2, fig.height=1.1}
rbind(results231, results232) %>% 
  ggplot(aes(x=frac_tau_h, y=kappa_coord, color=homeo, group=interaction(homeo, frac_tau_h))) +
  geom_boxplot() +
  scale_color_manual(values=c("black", "grey"), name="") 
```