### 2019-04-21 resistingrhythm/analysis/paper_figures_v4.Rmd
Fig tweaks

---
title: "Analysis : exp220 - 228 (Nov 2018, for J Neurophys draft 2018)"
output: html_notebook
---

```{r, message=FALSE}
library(ggplot2)
library(latex2exp)
library(dplyr)
library(magrittr)
library(readr)
library(gridExtra)
library(grid)
library(ggpubr)
library(cowplot)

data_path <- "~/Code/resistingrhythm/data/"
```

# Load all data
```{r, message=FALSE}
# -----------------------------------------------------------------------
# 16: Burst E, with H
exp_code <- 224
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130, 135, 140, 145, 150, 155, 160)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0)
stim_codes <- c(3, 4) # 4

results224 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    results224 <- rbind(results224, df)
  }
}

results224 %>%
  filter(rate_k > 0) -> results224

# -----------------------------------------------------------------------
# 15: Sustained E, with H
exp_code <- 221
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130, 135, 140, 145, 150, 155, 160)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0)
stim_codes <- c(3, 4)

results221 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$stim_code <- rep(s_c, nrow(df))

    results221 <- rbind(results221, df)
  }
}

results221 %>%
  filter(rate_k > 0) -> results221

# -----------------------------------------------------------------------
# 22: Sustained E, with NO H
exp_code <- 220
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130, 135, 140, 145, 150, 155, 160)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0)
stim_codes <- c(3, 4)

results220 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))

    results220 <- rbind(results220, df)
  }
}

results220 %>%
  filter(rate_k > 0) -> results220

# -----------------------------------------------------------------------
# 21: Sustained I, with H
exp_code <- 223
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130, 135, 140, 145, 150, 155, 160)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0)
stim_codes <- c(3, 4)

results223 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("GABA", nrow(df))

    results223 <- rbind(results223, df)
  }
}

results223 %>%
  filter(rate_k > 0) -> results223

# -----------------------------------------------------------------------
# 23: Sustained I, with NO H
exp_code <- 222
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130, 135, 140, 145, 150, 155, 160)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0)
stim_codes <- c(3, 4)

results222 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("GABA", nrow(df))

    results222 <- rbind(results222, df)
  }
}

results222 %>%
  filter(rate_k > 0) -> results222

# -----------------------------------------------------------------------
# 30: Burst I, with H
# HERE
exp_code <- 227
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130, 135, 140, 145, 150, 155, 160)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0)
stim_codes <- c(3, 4)

results227 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("GABA", nrow(df))

    results227 <- rbind(results227, df)
  }
}

results227 %>%
  filter(rate_k > 0) -> results227

# -----------------------------------------------------------------------
# Calcium exp, E, sustained
exp_code <- 228
osc_codes <- c(105, 115, 125, 135, 145, 155)
osc_rates <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5)
stim_codes <- c(3)

ca_targets <- c(0.0026, 0.0028, 0.0032, 0.0034)

results228 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    for(ca in ca_targets){
      o_c <- osc_codes[i]
      o_r <- osc_rates[i]

      # Load the csv
      df <- read_csv(paste(data_path,
                           "exp", as.character(exp_code),
                           "_stim", as.character(s_c),
                           "_osc", as.character(o_c),
                           "_ca", as.character(ca),
                           ".csv", sep=""))

      # Add metadata
      df$osc_code <- rep(o_c, nrow(df))
      df$osc_rate <- rep(o_r, nrow(df))
      df$stim_code <- rep(s_c, nrow(df))
      df$Ca_target <- rep(ca, nrow(df))
      df$synapse <- rep("AMPA", nrow(df))

      results228 <- rbind(results228, df)
    }
  }
}

# -----------------------------------------------------------------------
# # Add default Ca value, from exp
exp_code <- 221
osc_codes <- c(105, 115, 125, 135, 145, 155)
osc_rates <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5)
stim_codes <- c(3)
ca_targets <- c(0.003, 0.003, 0.003, 0.003)

df <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    for(ca in ca_targets){

      o_c <- osc_codes[i]
      o_r <- osc_rates[i]

      # Load the csv
      df <- read_csv(paste(data_path,
                           "exp", as.character(exp_code),
                           "_stim", as.character(s_c),
                           "_osc", as.character(o_c),
                           ".csv", sep=""))

      # Add metadata
      df$osc_code <- rep(o_c, nrow(df))
      df$osc_rate <- rep(o_r, nrow(df))
      df$stim_code <- rep(s_c, nrow(df))
      df$Ca_target <- rep(ca, nrow(df))
      df$synapse <- rep("AMPA", nrow(df))

      results228 <- rbind(results228, df)
    }
  }
}

# Clean up
rm(df, exp_code, i, o_c, o_r, osc_codes, osc_rates, s_c, stim_codes)

# 231
tau_h <- 600
exp_code <- 231
n_cycles <- c("4", "8", "480")

results231 <- NULL
for(i in 1:length(n_cycles)){
  n_c <- n_cycles[i]

  # Load the csv
  df <- read_csv(paste(data_path, 
                       "exp", as.character(exp_code), 
                       "_c", n_c, 
                       ".csv", sep=""))  
  # Add metadata
  df$n_cycles <- rep(n_c, nrow(df))
  df$frac_tau_h <- rep(as.character(
    (
      round(as.numeric(n_c)*(1/8)/tau_h, digits=3)
    ), nrow(df)
    )
  )
  df$osc_rate <- rep("6", nrow(df))
  df$freq <- rep("b", nrow(df))
  df$synapse <- rep("Excitatory", nrow(df))
  df$homeo <- rep("Homeostasis", nrow(df))
  if(n_c <= 8){
    df$mode <- rep("Burst", nrow(df))
  } else {
    df$mode <- rep("Tonic", nrow(df))
  }

  results231 <- rbind(results231, df)
}

# Realistic tau_h and oscillation.
exp_code <- 232
results232 <- NULL
for(i in 1:length(n_cycles)){
  n_c <- n_cycles[i]

  # Load the csv
  df <- read_csv(paste(data_path, 
                       "exp", as.character(exp_code), 
                       "_c", n_c, 
                       ".csv", sep=""))  
  # Add metadata
  df$n_cycles <- rep(n_c, nrow(df))
  df$frac_tau_h <- rep(as.character(
    (
      round(as.numeric(n_c)*(1/8)/tau_h, digits=3)
    ), nrow(df)
    )
  )
  df$osc_rate <- rep("6", nrow(df))
  df$freq <- rep("b", nrow(df))
  df$synapse <- rep("Excitatory", nrow(df))
  df$homeo <- rep("No homeostasis", nrow(df))
  if(n_c <= 8){
    df$mode <- rep("Burst", nrow(df))
  } else {
    df$mode <- rep("Tonic", nrow(df))
  }

  results232 <- rbind(results232, df)
}

rm(df, n_c, i, exp_code)
```

# Figure: Main results
```{r fig.height=2.25, fig.width=6}
# Kappa no H, E
bind_rows(results220, results222) %>%
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>% 
  group_by(synapse, osc_rate) %>% 
  mutate(M=mean(kappa_coord), SD=sd(kappa_coord)) %>% 
  ggplot(mapping=aes(x=osc_rate, y=M, group=osc_rate, color=synapse)) +
    geom_point(size=1.25) +
    geom_errorbar(mapping=aes(x=osc_rate, ymin=M - SD, ymax=M + SD, color=synapse),
                  width=.15, size=0.2) +
    scale_color_manual(values=c("black", "grey")) +
    lims(y=c(0.0, 0.2)) +
    labs(title="b.", y="Kappa", x="Osc. rate") +
    theme_classic2() +
    theme(plot.title = element_text(hjust = -.3, size=18, face="bold"),
          legend.position="none") -> p1

# Kappa with H, E
bind_rows(results221, results223) %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(synapse, osc_rate) %>% 
  mutate(M=mean(kappa_coord), SD=sd(kappa_coord)) %>% 
  ggplot(mapping=aes(x=osc_rate, y=M, group=osc_rate, color=synapse)) +
    geom_point(size=1.25) +
    geom_errorbar(mapping=aes(x=osc_rate, ymin=M - SD, ymax=M + SD, color=synapse),
                  width=.15, size=0.2) +
    scale_color_manual(values=c("black", "grey")) +
    lims(y=c(0.0, 0.2)) +
    labs(title="d.", y="Kappa", x="Osc. rate") +
    theme_classic2() +
    theme(plot.title = element_text(hjust = -.3, size=18, face="bold"),
          legend.position="none") -> p2

# Rate no H, E
bind_rows(results220, results222) %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>% 
  group_by(synapse, osc_rate) %>% 
  mutate(M=mean(rate_k), SD=sd(rate_k)) %>% 
  ggplot(mapping=aes(x=osc_rate, y=M, group=osc_rate, color=synapse)) +
  geom_point(size=1.25) +
  geom_errorbar(mapping=aes(x=osc_rate, ymin=M - SD, ymax=M + SD, color=synapse),
                width=.15, size=0.2) +
  scale_color_manual(values=c("black", "grey")) +
  labs(title="a.", y="Pop. rate", x="Osc. rate") +
  lims(y=c(0.0, 180.0)) +
  guides(fill = guide_legend(override.aes = list(shape = 15, size = 7, alpha=0.8))) +
  theme_classic2() +
  theme(plot.title = element_text(hjust = -.3, size=18, face="bold"),
        legend.position="none") -> p3

# -
# Rate, with H, E
bind_rows(results221, results223) %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(synapse, osc_rate) %>% 
  mutate(M=mean(rate_k), SD=sd(rate_k)) %>% 
  ggplot(mapping=aes(x=osc_rate, y=M, group=osc_rate, color=synapse)) +
  geom_point(size=1.25) +
  geom_errorbar(mapping=aes(x=osc_rate, ymin=M - SD, ymax=M + SD, color=synapse), 
                width=.15, size=0.2) +
  scale_color_manual(values=c("black", "grey"), name="Tonic") +
  labs(title="c.", y="Pop. rate", x="Osc. rate") +
  lims(y=c(0.0, 180.0)) +
  theme_classic2() +
  theme(
    legend.position = c(-.5, -0.2),
    legend.background = element_rect(colour ="black"),
    legend.key.size = unit(.5, "cm"),
    plot.title = element_text(hjust = -.3, size=18, face="bold")) -> p4

# --------------------------------------------------------
bind_rows(results224, results227)  %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(synapse, osc_rate) %>% 
  mutate(M=mean(kappa_coord), SD=sd(kappa_coord)) %>% 
  ggplot(mapping=aes(x=osc_rate, y=M, group=osc_rate, color=synapse)) +
    geom_point(size=1.25) +
    geom_errorbar(mapping=aes(x=osc_rate, ymin=M - SD, ymax=M + SD, color=synapse), 
                width=.15, size=0.2) +
    scale_color_manual(values=c("darkgoldenrod3", "goldenrod1")) +
    lims(y=c(0.0, 0.2)) +
    labs(title="f.", y="Kappa", x="Osc. rate") +
    theme_classic2() +
    theme(plot.title = element_text(hjust = -.3, size=18, face="bold"),
          legend.position="none") -> p5

# -
bind_rows(results224, results227)  %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(synapse, osc_rate) %>% 
  mutate(M=mean(rate_k), SD=sd(rate_k)) %>% 
  ggplot(mapping=aes(x=osc_rate, y=M, group=osc_rate, color=synapse)) +
    geom_point(size=1.25) +
    geom_errorbar(mapping=aes(x=osc_rate, ymin=M - SD, ymax=M + SD, color=synapse), 
                width=.15, size=0.2) +
    scale_color_manual(values=c("darkgoldenrod3", "goldenrod1"), name="Burst") +
    labs(title="e.", y="Pop. rate", x="Osc. rate") +
    lims(y=c(0.0, 180.0)) +
    theme_classic2() +
  theme(
    legend.position = c(1.5, -1.25),
    legend.title = element_text(colour = 'darkgoldenrod3'),
    legend.background = element_rect(colour ="black"),
    legend.key.size = unit(.5, "cm"),
    plot.title = element_text(hjust = -.3, size=18, face="bold")) -> p6

# -
tmp3 <- results221
tmp3$osc_mode <- rep("Tonic", nrow(tmp3))
tmp4 <- results224
tmp4$osc_mode <- rep("Burst", nrow(tmp4))

tmp5 <- results223
tmp5$osc_mode <- rep("Tonic", nrow(tmp5))
tmp6 <- results227
tmp6$osc_mode <- rep("Burst", nrow(tmp6))

bind_rows(tmp3, tmp4, tmp5, tmp6) %>% 
  # filter(synapse=="AMPA") %>% 
  ggplot(aes(x=osc_mode, y=rate_k, color=osc_mode)) +
  geom_point(size=-1, aes(fill=osc_mode)) +
  geom_boxplot(notch = TRUE, width=0.4,  outlier.size=0.05, outlier.alpha=0.5, show.legend=FALSE) +
  facet_grid(.~synapse) +
  labs(title="g.", y="Pop. rate", x="") +
  theme_classic2() +
  scale_fill_manual(values=c("black", "black"), name="Oscillation") +
  scale_color_manual(values=c("black", "black"), name="Oscillation") +
  stat_compare_means(method="wilcox.test", label="p.signif", 
                     label.x=1.3, size=3, show.legend=FALSE) +
  # lims(y=c(0.0, 180.0)) +
  guides(fill = guide_legend(override.aes = list(shape = 15, size = 7, alpha=0.8))) +
  theme(plot.title = element_text(hjust = -.1, size=18, face="bold"), 
          legend.position="none",
          strip.background = element_blank(),
          axis.text.x = element_text(angle = 70, hjust = 1)) -> p7

# -
p8 <- textGrob("No homeostatsis")
p9 <- textGrob("Homeostatsis")
p10 <- textGrob("Synchrony")
p11 <- textGrob("Excitability")

# -
layout_matrix <- rbind(
  c(NA, 8, 8, 8, NA, 9, 9, 9, 9, 9, 9, NA, NA),
  c(11, 3, 3, 3, NA, 4, 4, 4, 6, 6, 6, 7, 7),
  c(11, 3, 3, 3, NA, 4, 4, 4, 6, 6, 6, 7, 7),
  c(11, 3, 3, 3, NA, 4, 4, 4, 6, 6, 6, 7, 7),
  c(11, 3, 3, 3, NA, 4, 4, 4, 6, 6, 6, 7, 7),
  c(10, 1, 1, 1, NA, 2, 2, 2, 5, 5, 5, 7, 7), 
  c(10, 1, 1, 1, NA, 2, 2, 2, 5, 5, 5, NA, NA), 
  c(10, 1, 1, 1, NA, 2, 2, 2, 5, 5, 5, NA, NA), 
  c(10, 1, 1, 1, NA, 2, 2, 2, 5, 5, 5, NA, NA)
)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, layout_matrix=layout_matrix)

# -
png(file="fig2.png", bg="white", width=12.0, height=4.5, unit="in", res=300)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, layout_matrix=layout_matrix)
dev.off()
```

```{r}
print(wilcox.test(x=tmp3$rate_k, y=tmp4$rate_k))
```

# Figure: Error
```{r, fig.width=3.2, fig.height=1.2}
# ------------------------------------------------------------
# Kappa
results220 %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(osc_rate) %>% 
  mutate(
    M_error=mean(mse_error), SD_error=sd(mse_error), 
    M_rate_k=mean(rate_k), SD_rate_k=sd(rate_k), 
    M_kappa=mean(kappa_coord), SD_kappa=sd(kappa_coord)) -> tmp1
tmp1$mode <- rep("No homeostatsis", nrow(tmp1))
results221 %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(osc_rate) %>% 
  mutate(
    M_error=mean(mse_error), SD_error=sd(mse_error), 
    M_rate_k=mean(rate_k), SD_rate_k=sd(rate_k), 
    M_kappa=mean(kappa_coord), SD_kappa=sd(kappa_coord)) -> tmp2
tmp2$mode <- rep("Homeostatsis", nrow(tmp2))

# Kappa
bind_rows(tmp1, tmp2) %>% 
ggplot(mapping=aes(x=M_kappa, y=M_error, color=mode)) +
  geom_point(size=1.25) +
  geom_errorbar(mapping=aes(x=M_kappa, ymin=M_error - SD_error, ymax=M_error + SD_error),
                width=0.004, size=0.2) +
  geom_errorbarh(mapping=aes(y=M_error, xmin=M_kappa - SD_kappa, xmax=M_kappa + SD_kappa),
                 height=0.002, size=0.2) +
  scale_color_manual(values=c("black", "darkgrey")) +
  labs(title="a.", y="Log error (s)", x="Kappa") +
  lims(x=c(0.0, 0.25)) +
  scale_y_log10() + annotation_logticks() +
  theme_classic2() + 
  theme(plot.title = element_text(hjust = -.3, size=18, face="bold"),
        legend.position = "none") -> p1

# Rate
# bind_rows(tmp1, tmp2) %>% 
# ggplot(mapping=aes(x=M_rate_k, y=M_error, color=mode)) +
#   geom_point(size=0.2) +
#   geom_errorbar(mapping=aes(x=M_rate_k, ymin=M_error - SD_error, ymax=M_error + SD_error),
#                 width=0.004, size=0.2) +
#   geom_errorbarh(mapping=aes(y=M_error, xmin=M_rate_k - SD_rate_k, xmax=M_rate_k + SD_rate_k),
#                  height=0.002, size=0.2) +
#   scale_color_manual(values=c("black", "darkgrey")) +
#   labs(title="c.", y="Log error (s)", x="Pop. rate") +
#   # lims(x=c(0.0, 0.25)) +
#   scale_y_log10() + annotation_logticks() +
#   theme_classic2() + 
#   theme(plot.title = element_text(hjust = -.3, size=18, face="bold"),
#         legend.position = "none") -> p3

# ---------------------------------------------------------------------
results220 %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(osc_rate) %>% 
  mutate(
      M_error=mean(mse_error), SD_error=sd(mse_error), 
      M_rate_k=mean(rate_k), SD_rate_k=sd(rate_k), 
      M_kappa=mean(kappa_coord), SD_kappa=sd(kappa_coord)) -> tmp1
tmp1$mode <- rep("No homeostatsis", nrow(tmp1))
results224 %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(osc_rate) %>% 
  mutate(
    M_error=mean(mse_error), SD_error=sd(mse_error), 
    M_rate_k=mean(rate_k), SD_rate_k=sd(rate_k), 
    M_kappa=mean(kappa_coord), SD_kappa=sd(kappa_coord)) -> tmp2
tmp2$mode <- rep("Homeostatsis", nrow(tmp2))

# Kappa
bind_rows(tmp1, tmp2) %>% 
ggplot(mapping=aes(x=M_kappa, y=M_error, color=mode)) +
  geom_point(size=1.25) +
  geom_errorbar(mapping=aes(x=M_kappa, ymin=M_error - SD_error, ymax=M_error + SD_error),
                width=0.004, size=0.2) +
  geom_errorbarh(mapping=aes(y=M_error, xmin=M_kappa - SD_kappa, xmax=M_kappa + SD_kappa),
                 height=0.002, size=0.2) +
  scale_color_manual(values=c("black", "darkgrey")) +
  labs(title="b.", y="Log error (s)", x="Kappa") +
  lims(x=c(0.0, 0.25)) + annotation_logticks() +
  theme_classic2() + 
  scale_y_log10() +
  theme(plot.title = element_text(hjust = -.3, size=18, face="bold"),
        legend.title=element_blank()) -> p2

# Rate
# bind_rows(tmp1, tmp2) %>% 
# ggplot(mapping=aes(x=M_rate_k, y=M_error, color=mode)) +
#   geom_point(size=0.2) +
#   geom_errorbar(mapping=aes(x=M_rate_k, ymin=M_error - SD_error, ymax=M_error + SD_error),
#                 width=0.004, size=0.2) +
#   geom_errorbarh(mapping=aes(y=M_error, xmin=M_rate_k - SD_rate_k, xmax=M_rate_k + SD_rate_k),
#                  height=0.002, size=0.2) +
#   scale_color_manual(values=c("black", "darkgrey")) +
#   labs(title="d.", y="Log error (s)", x="Pop. rate") +
#   # lims(x=c(0.0, 0.25)) +
#   theme_classic2() + 
#   scale_y_log10() +
#   theme(plot.title = element_text(hjust = -.3, size=18, face="bold"),
#         legend.title=element_blank()) -> p4

# -
p5 <- textGrob("Tonic")
p6 <- textGrob("Burst", gp=gpar(col="darkgoldenrod3"))

# -
layout_matrix <- rbind(
  c(3, 3, 3, 4, 4, 4, 4, NA),
  c(1, 1, 1, 2, 2, 2, 2, 2),
  c(1, 1, 1, 2, 2, 2, 2, 2),
  c(1, 1, 1, 2, 2, 2, 2, 2),
  c(1, 1, 1, 2, 2, 2, 2, 2)
)
grid.arrange(p1, p2, p5, p6, layout_matrix=layout_matrix)

png(file="fig3.png", bg="white", width=6.4, height=2.4, unit="in", res=300)
grid.arrange(p1, p2, p5, p6, layout_matrix=layout_matrix)
dev.off()
```

# Figure: Control
```{r, fig.width=2.4, fig.height=2.8}
# -
results228 %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(osc_rate, Ca_target) %>% 
  summarise(M=mean(kappa_coord), SD=sd(kappa_coord)) -> tmp1

tmp1 %>% 
  ggplot(aes(x=osc_rate, y=M, color=factor(Ca_target, 
                                           levels = c(0.0026, 0.0028, 0.003, 0.0032, 0.0034)))) +
  geom_line(size=0.8) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin=M - SD, ymax=M + SD), alpha=0.5, width=.1, size=0.4)  +
  labs(title="b.", y="Kappa", x="Osc. rate") +
  scale_color_manual(
    values=c("plum1", "plum2", "black", "plum3", "plum4"), 
    guide=FALSE) +
  theme_classic2() +
  theme(plot.title = element_text(hjust = -.5, size=18, face="bold")) -> p1

# -
results228 %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>%
  group_by(osc_rate, Ca_target) %>% 
  summarise(M=mean(rate_k), SD=sd(rate_k)) -> tmp2

tmp2 %>% 
ggplot(aes(x=osc_rate, y=M, color=factor(Ca_target, 
                                         levels = c(0.0026, 0.0028, 0.003, 0.0032, 0.0034)))) +
  geom_line(size=0.8) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin=M - SD, ymax=M + SD), alpha=0.5, width=.1, size=0.4)  +
  labs(title="a.", y="Pop. rate", x="Osc. rate") +
  scale_color_manual(
    values=c("plum1", "plum2", "black", "plum3", "plum4"), 
    name="Target [Ca]\n(mmolar)") +
  theme_classic2() +
  theme(
    plot.title = element_text(hjust = -.5, size=18, face="bold"),
    legend.position = c(1.35, -0.5),
    legend.background = element_rect(colour ="black"),
    legend.key.size = unit(.4, "cm")) -> p2

# -----------------------------------------------------------------------------------
# Frac of tau_h
# rate
rbind(results231, results232) %>% 
  ggplot(aes(x=frac_tau_h, y=rate_k, color=homeo, group=interaction(homeo, frac_tau_h))) +
  geom_boxplot(size=0.6) +
  stat_compare_means(method="wilcox.test", label="p.signif", 
                     label.x=1.3, size=3, show.legend=FALSE) +
  labs(title="c.", y="Pop. rate", x=TeX("$t_{oscillation} / \\tau_{h}$")) +
  scale_color_manual(values=c("black", "grey"), name=NA) +
  theme_classic2() +
  theme(
    plot.title = element_text(hjust = -.5, size=18, face="bold"),
    legend.background = element_rect(colour ="black"),
    legend.position = c(1.35, 0.5),
    legend.title=element_blank(),
    legend.key.size = unit(.3, "cm")) -> p6

# -
p3 <- textGrob("")
p4 <- textGrob("Target [Ca]\nperturbation")
p8 <- textGrob(TeX("Biological $\\tau_h$"))
p5 <- nullGrob()
p9 <- nullGrob()

# -
# layout_matrix <- rbind(
#   c(4, 2, 2, 5), 6, 6, 5),
#   c(4, 2, 2, 5), 6, 6, 5),
#   c(4, 2, 2, 5), 6, 6, 5),
#   c(4, 2, 2, 5), 6, 6, 5),
#   c(3, 1, 1, 5), 7, 7, 5),
#   c(3, 1, 1, 5), 7, 7, 5),
#   c(3, 1, 1, 5), 7, 7, 5), 
#   c(3, 1, 1, 5), 7, 7, 5) 
# )

layout_matrix <- rbind(
  c(4, 2, 2, 5), 
  c(4, 2, 2, 5), 
  c(4, 2, 2, 5), 
  c(4, 2, 2, 5), 
  c(3, 1, 1, 5), 
  c(3, 1, 1, 5), 
  c(3, 1, 1, 5),  
  c(3, 1, 1, 9), 
  c(8, 6, 6, 9),
  c(8, 6, 6, 9),
  c(8, 6, 6, 9),
  c(8, 6, 6, 9)
)
grid.arrange(p1, p2, p3, p4, p5, p6, p8, p9, layout_matrix=layout_matrix)

png(file="fig4.png", bg="white", width=4.8, height=5.6, unit="in", res=300)
grid.arrange(p1, p2, p3, p4, p5, p6, p8, p9, layout_matrix=layout_matrix)
dev.off()
```