### 2019-05-28 resistingrhythm/analysis/exp244-245.Rmd
Analyze 244/45

---
title: "Analysis : exp244-245 analysis"
output: html_notebook
---

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)
library(readr)
library(gridExtra)
library(grid)
library(ggpubr)
library(cowplot)

data_path <- "~/Code/resistingrhythm/data/"
```

# Load all data
```{r, message=FALSE}
# -----------------------------------------------------------------------
# LFP E, no H
exp_code <- 244
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4) # 4

results244 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    results244 <- rbind(results244, df)
  }
}

results244 %>%
  filter(rate_k > 0) -> results244

# LFP E, w/ H
exp_code <- 245
osc_codes <- c(100, 101, 102, 103, 104, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4) # 4

results245 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    results245 <- rbind(results245, df)
  }
}

results245 %>%
  filter(rate_k > 0) -> results245

results <- rbind(results244, results245)
rm(df)
```

```{r}
results %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>% 
  ggplot(aes(x=osc_rate, y=kappa_coord, group=interaction(osc_rate, homeo), color=homeo)) +
  geom_boxplot() +
  theme_classic() +
  scale_color_grey()
```

```{r}
results %>% 
  filter(osc_rate < 0.1 | osc_rate > 0.4) %>% 
  mutate(kappa_coord=replace(kappa_coord, rate_k < 1, NA)) %>% 
  ggplot(aes(x=osc_rate, y=rate_k, group=interaction(osc_rate, homeo), color=homeo)) +
  geom_boxplot() +
  theme_classic() +
  scale_color_grey()
```