### 2019-05-28 resistingrhythm/analysis/exp249-253.Rmd
Analyze 249-253

---
title: "Analysis : exp249-253 analysis"
output: html_notebook
---

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)
library(readr)
library(gridExtra)
library(grid)
library(ggpubr)
library(cowplot)

data_path <- "~/Code/resistingrhythm/data/"
```

# Load all data

- Explore gamma=60Hz, gamma=90Hz, ripple=150 Hz

```{r, message=FALSE}
# ----------------------------------------------------------------------
# Load 8 Hz theta AMPA
# 16: Burst E, with H
exp_code <- 224
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4) # 4

results224 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    df$freq <- rep(8, nrow(df))
    df$freq_name <- rep("Theta (8 Hz)", nrow(df))
    results224 <- rbind(results224, df)
  }
}
results224 %>%
  filter(rate_k > 0) -> results224

# 22: Burst E, NO H @ 8 hz
exp_code <- 220
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results220 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    df$freq <- rep(8, nrow(df))
    df$freq_name <- rep("Theta (8 Hz)", nrow(df))
    results220 <- rbind(results220, df)
  }
}
results220 %>%
  filter(rate_k > 0) -> results220

# ----------------------------------------------------------------------
# 60 Hz gamma
# exp248
exp_code <- 248
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results248 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    df$freq <- rep(60, nrow(df))
    df$freq_name <- rep("Slow gamma (60 Hz)", nrow(df))
    results248 <- rbind(results248, df)
  }
}
results248 %>%
  filter(rate_k > 0) -> results248

# exp249, no H
exp_code <- 249
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results249 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(60, nrow(df))
    df$freq_name <- rep("Slow gamma (60 Hz)", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    results249 <- rbind(results249, df)
  }
}
results249 %>%
  filter(rate_k > 0) -> results249

# ----------------------------------------------------------------------
# 90 Hz gmma
# exp250, w H
exp_code <- 250
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results250 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(90, nrow(df))
    df$freq_name <- rep("Fast gamma (90 Hz)", nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    results250 <- rbind(results250, df)
  }
}
results250 %>%
  filter(rate_k > 0) -> results250

# exp251, no H
exp_code <- 251
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results251 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(90, nrow(df))
    df$freq_name <- rep("Fast gamma (90 Hz)", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    results251 <- rbind(results251, df)
  }
}
results251 %>%
  filter(rate_k > 0) -> results251

# ----------------------------------------------------------------------
# 150 Hz ripple

# exp252 w/ H
exp_code <- 252
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results252 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(150, nrow(df))
    df$freq_name <- rep("Ripple (150 Hz)", nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    results252 <- rbind(results252, df)
  }
}
results252 %>%
  filter(rate_k > 0) -> results252

# exp253, no H
exp_code <- 253
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results253 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_burst", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(150, nrow(df))
    df$freq_name <- rep("Ripple (150 Hz)", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    results253 <- rbind(results253, df)
  }
}
results253 %>%
  filter(rate_k > 0) -> results253

results <- rbind(
  results220, results224, results248, results249, results250, results251, results252, results253)
results$freq_name <- factor(results$freq_name, 
                               levels=c("Theta (8 Hz)", "Slow gamma (60 Hz)", "Fast gamma (90 Hz)", "Ripple (150 Hz)"))
```

```{r}
results %>% 
  ggplot(aes(x=osc_rate, y=kappa_coord, group=interaction(osc_rate, homeo), color=homeo)) +
  geom_boxplot() +
  theme_classic() +
  scale_color_grey() +
  facet_grid(.~freq_name, scales = "free")
```

```{r, fig.width=4, fig.height=1.2}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k, group=interaction(osc_rate, homeo), color=homeo)) +
  geom_boxplot() +
  theme_classic() +
  scale_color_grey() + 
  facet_grid(.~freq_name, scales = "free")
```