### 2019-05-28 resistingrhythm/analysis/exp254-267.Rmd
Add analysis for exp254-275

---
title: "Analysis : exp249-253 analysis"
output: html_notebook
---

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)
library(readr)
library(gridExtra)
library(grid)
library(ggpubr)
library(cowplot)

data_path <- "~/Code/resistingrhythm/data/"
```

# Load all data

- Explore gamma=60Hz, gamma=90Hz, ripple=150 Hz

```{r, message=FALSE}
# ----------------------------------------------------------------------
# Load 8 Hz theta AMPA
# 16: Burst E, with H
exp_code <- 221
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4) # 4

results221 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    df$freq <- rep(8, nrow(df))
    results221 <- rbind(results221, df)
  }
}
results221 %>%
  filter(rate_k > 0) -> results221

# 22: Burst E, NO H @ 8 hz
exp_code <- 220
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results220 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_osc", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    df$freq <- rep(8, nrow(df))
    results220 <- rbind(results220, df)
  }
}
results220 %>%
  filter(rate_k > 0) -> results220

# ----------------------------------------------------------------------
# @ 6 Hz
#
# No H
exp_code <- 254
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results254 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    df$freq <- rep(6, nrow(df))
    results254 <- rbind(results254, df)
  }
}
results254 %>%
  filter(rate_k > 0) -> results254

# H
exp_code <- 255
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results255 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(6, nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    results255 <- rbind(results255, df)
  }
}
results255 %>%
  filter(rate_k > 0) -> results255

# ----------------------------------------------------------------------
# @ 7 Hz
#
# No H
exp_code <- 258
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results258 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    df$freq <- rep(7, nrow(df))
    results258 <- rbind(results258, df)
  }
}
results258 %>%
  filter(rate_k > 0) -> results258

# H
exp_code <- 259
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results259 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(7, nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    results259 <- rbind(results259, df)
  }
}
results259 %>%
  filter(rate_k > 0) -> results259

# ----------------------------------------------------------------------
# @ 9 Hz
#
# No H
exp_code <- 262
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results262 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    df$freq <- rep(9, nrow(df))
    results262 <- rbind(results262, df)
  }
}
results262 %>%
  filter(rate_k > 0) -> results262

# H
exp_code <- 263
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results263 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(9, nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    results263 <- rbind(results263, df)
  }
}
results263 %>%
  filter(rate_k > 0) -> results263

# ----------------------------------------------------------------------
# @ 10 Hz
#
# No H
exp_code <- 266
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results266 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$homeo <- rep("No homeostasis", nrow(df))
    df$freq <- rep(10, nrow(df))
    results266 <- rbind(results266, df)
  }
}
results266 %>%
  filter(rate_k > 0) -> results266

# H
exp_code <- 267
osc_codes <- c(100, 105, 110, 115, 120, 125, 130)
osc_rates <- c(0, 0.5, 1.0, 1.5, 2.0, 2.5, 3.0)
stim_codes <- c(3, 4)

results267 <- NULL
for(i in 1:length(osc_codes)){
  for(s_c in stim_codes){
    o_c <- osc_codes[i]
    o_r <- osc_rates[i]

    # Load the csv
    df <- read_csv(paste(data_path, 
                         "exp", as.character(exp_code), 
                         "_stim", as.character(s_c), 
                         "_lfp", as.character(o_c),
                         ".csv", sep=""))

    # Add metadata
    df$osc_code <- rep(o_c, nrow(df))
    df$osc_rate <- rep(o_r, nrow(df))
    df$stim_code <- rep(s_c, nrow(df))
    df$synapse <- rep("AMPA", nrow(df))
    df$freq <- rep(10, nrow(df))
    df$homeo <- rep("Homeostasis", nrow(df))
    results267 <- rbind(results267, df)
  }
}
results267 %>%
  filter(rate_k > 0) -> results267

results <- rbind(
  results220, results221, results254, results255, results258, results259, results262, results263, results266, results267)
```

```{r}
results %>% 
  ggplot(aes(x=osc_rate, y=kappa_coord, group=interaction(osc_rate, homeo), color=homeo)) +
  geom_boxplot() +
  theme_classic() +
  scale_color_grey() +
  facet_grid(.~freq, scales = "free")
```

```{r, fig.width=4, fig.height=1.2}
results %>% 
  ggplot(aes(x=osc_rate, y=rate_k, group=interaction(osc_rate, homeo), color=homeo)) +
  geom_boxplot() +
  theme_classic() +
  scale_color_grey() + 
  facet_grid(.~freq, scales = "free")
```